Tidy data structure to support exploration and modeling of temporal-context data

### 2018-10-30 main.Rmd
added a paragragh to delayed quantiles

Figure \@ref(fig:carrier-mosaic) visually presents the end product of aggregating the number of on-time and delayed flights to the year interval by carriers. This pipeline is initialized defining a new variable if the flight is delayed, and involves summarizing the tallies of on-time and delayed flights for each carrier annually. To prepare the summarized data for a mosaic plot, it is further manipulated by melting new tallies into a single column. The flow chart shown as Figure \@ref(fig:flight-pipeline1) demonstrates the operations undertaking in a data pipeline. The input to this pipeline is a tsibble of irregular interval, and the output ends up with a tsibble of unknown interval. The final data is each carrier along with a single year, in which the interval cannot be determined. It in turn feeds into the mosaic display. Note that Southwest Airlines, as the largest carrier, operates less efficiently compared to Delta, in Figure \@ref(fig:carrier-mosaic).

@@

A closer examination of New York airports will give an indication about how well the busiest airports manage the outflow traffic on a daily basis. A subset that contains observations for EWR, JFK and LGA airports is obtained first. The succeeding operations compute delayed percentages every day at each airport, which are framed as grey lines in Figure \@ref(fig:nyc-monthly-plot). LGA fluctuates a lot compared to the other two. What superimposes on these lines is two-month moving averages so that a temporal trend is more visible. The number of days for each month is variable. Moving averages for two months call for computing weighted mean. But this can also be accomplished using a pair of commonly used verbs--`nest()` and `unnest()` to handle list-columns, without specifying weights. The sliding of large window size smoothes out the fluctuations and gives a stable trend around 25% over the year.

@@

@@

The data has been transformed in a way that the variables could be mapped to graphical elements. Figure \@ref(fig:draw-qtl) shows the plot of the transformed data using the grammar of graphics. It suggests to leave early in a day and fly on Tuesdays in a week to avoid severe delay.

@@

(ref:draw-qtl-cap) Small multiples of lines about departure delay against time of day, faceted by day of week and quantiles. A blue horizontal line is drawn at 15-minute departure delay, which highlights that the departing flights who delay less than 15 minutes are considered as on time.

>>>>

Figure \@ref(fig:carrier-mosaic) visually presents the end product of aggregating the number of on-time and delayed flights to the year interval by carriers. This pipeline is initialized defining a new variable if the flight is delayed, and involves summarizing the tallies of on-time and delayed flights for each carrier annually. To prepare the summarized data for a mosaic plot, it is further manipulated by melting new tallies into a single column. The flow chart shown as Figure \@ref(fig:flight-pipeline1) demonstrates the operations undertaking in a data pipeline. The input to this pipeline is a tsibble of irregular interval, and the output ends up with a tsibble of unknown interval. The final data is each carrier along with a single year, thereby the interval undetermined. It in turn feeds into the mosaic display. Note that Southwest Airlines, as the largest carrier, operates less efficiently compared to Delta, in Figure \@ref(fig:carrier-mosaic).

@@

A closer examination of New York airports will give an indication about how well the busiest airports manage the outflow traffic on a daily basis. A subset that contains observations for EWR, JFK and LGA airports is obtained first. The succeeding operations compute delayed percentages every day at each airport, which are framed as grey lines in Figure \@ref(fig:nyc-monthly-plot). LGA fluctuates a lot compared to the other two. What superimposes on these lines is two-month moving averages so that a temporal trend is more visible. The number of days for each month is variable. Moving averages for two months call for computing weighted mean. But this can also be accomplished using a pair of commonly used verbs--`nest()` and `unnest()` to handle list-columns, without worrying weights specification. The sliding with large window size smoothes out the fluctuations and gives a stable trend around 25% over the year.

@@

What time of day and day of week should we travel to avoid suffering from horrible delay? Figure \@ref(fig:draw-qtl) plots hourly quantile estimates across day of week in the form of small multiples. The upper-tail delay behaviors are of primary interest, and hence 50%, 80% and 95% quantiles are shown. To reduce the likelihood of delay suffering, it is recommended to avoid the peak hour at 18. As moving towards the upper extremes, the variations considerably increase, making departure time unpredictable.

@@

@@

(ref:draw-qtl-cap) Small multiples of lines about departure delay against time of day, faceting day of week and 50%, 80% and 95% quantiles. A blue horizontal line indicates the 15-minute on-time standard to help grasp the delay severity. Passengers are apt to hold up around 18 during a day, and are recommended to travel early. The variations increase substantially as the upper tails.
